Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality

### 2024-11-04 paper/paper.qmd
fixed spelling mistakes

The team also built a Bayesian MLR using the same approach and formula for Trump. It is in section @sec-trumpmodel.

@@

The team has provided results for a Trump Bayesian MLR model. It is in section @sec-trumpmodelresults.

@@

According to Figure @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States. There are also a few states with predictions not visible on the map, we will describe those predictions below.

@@

Based on the figures in @sec-data, the team identified three sources of bias beyond sample size and state: methodology, the timing of polls relative to the election date, and poll scores. From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability. This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys. Additionally, using multiple communication methods provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias. Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.

From @fig-polldate-canidate-pct, the team observed a decline in third-party candidate support as the election approached. Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes. This trend may be worsened by decreased participation in polls further from the election day. As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates. Additionally, the withdrawal of a prominent third-party candidate, Robert F. Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic. However, we find the earlier reasons more compelling, as the data in Figure @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.

Finally, as illustrated in Figure @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls. It is important to note that the FiveThirtyEight poll score system is partly based on historical accuracy. Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies. This potential undervaluation of Trump’s support could benefit the Harris campaign, by providing a warning to not repeat previous election mistakes.

@@

Also, each state has a limited number of polls available. While the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states having only 27. Moreover, as noted in section @sec-prediction, the analysis focused solely on contested races, specifically swing states. Therefore, many states were assigned to a candidate based on strong historical trends and not from predictions. Although this approach is reasonable, it may overlook current trends that could potentially shift a state's vote from Republican to Democrat or vice versa.

>>>>

The team also built a Bayesian MLR using the same approach and formula for Trump. It is in @sec-trumpmodel.

@@

The team has provided results for a Trump Bayesian MLR model. It is in @sec-trumpmodelresults.

@@

According to @fig-currentstateofelection, Kamala Harris is predicted to be the 47th President of the United States. There are also a few states with predictions not visible on the map, we will describe those predictions below.

@@

Based on the figures in @sec-data, the team identified three sources of bias beyond sample size and state: methodology, the timing of polls relative to the election date, and poll scores. From the analysis in @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability. This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys. Additionally, using multiple communication methods provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias. Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.

From @fig-polldate-canidate-pct, the team observed a decline in third-party candidate support as the election approached. Early in the election cycle, voters may entertain the idea of a viable third-party candidate, but as the election nears, interest in third-party options diminishes. This trend may be worsened by decreased participation in polls further from the election day. As a result, the people who do participate in those earlier polls are likely to be more informed and passionate, which are qualities that may be associated with support for third-party candidates. Additionally, the withdrawal of a prominent third-party candidate, Robert F. Kennedy Jr., on August 23, 2024 (74 days from the election), could influence this dynamic. However, we find the earlier reasons more compelling, as the data in @fig-polldate-canidate-pct do not indicate a significant drop in third-party support at that time, but rather a gradual decline.

Finally, as illustrated in @fig-pollscore-candidate-pct, more reliable polls tend to show a bias favoring Trump over Harris—not necessarily indicating a guaranteed victory for Trump, but rather suggesting he may outperform expectations in many polls. It is important to note that the FiveThirtyEight poll score system is partly based on historical accuracy. Given Trump's overperformance in the polls during the 2016 and 2020 elections, this scoring may be overcorrecting for past inaccuracies. This potential undervaluation of Trump’s support could benefit the Harris campaign, by providing a warning to not repeat previous election mistakes.

@@

Also, each state has a limited number of polls available. While the overall analysis considered a substantial number of polls, individual states often had fewer than 20 polls, with even the largest states having only 27. Moreover, as noted in @sec-prediction, the analysis focused solely on contested races, specifically swing states. Therefore, many states were assigned to a candidate based on strong historical trends and not from predictions. Although this approach is reasonable, it may overlook current trends that could potentially shift a state's vote from Republican to Democrat or vice versa.
